---
title: "STA 308 Final"
author: "Bailey Burdick"
date: "5/7/2022"
output: html_document
---
### This assignment is my STA 308 Final assignment. 
### The purpose of this assignment is to calculate the difference in unemployment between March 2021 and March 2022.
### Then, I am calculating the mean, standard deviation, and coefficient of variation for this difference.
### Lastly I put this information into a table. 


#### Importing tidyverse and the data sets
```{r}
library(tidyverse)
march_unemploy_data <- read.csv("https://tjfisher19.github.io/data/marchStateUnemployment.csv")
abb_codes_data <-read_csv("https://tjfisher19.github.io/data/state_abb_codes.csv")
census_regions_data <- read_csv("https://tjfisher19.github.io/data/censusRegions.csv")
```

#### Filtering out DC from all the data sets
```{r}
march_unemploy_50 <- march_unemploy_data %>% 
    filter(State != "District of Columbia") 

abb_codes_50 <- abb_codes_data %>% 
    filter(State != "District of Columbia") 

census_regions_50 <- census_regions_data %>% 
    filter(State != "DC") 
```

#### Merging the data sets together into "unemployment_abb"
```{r}
unemployment_abb<- merge(abb_codes_50 ,march_unemploy_50, by = "State" )

unemployment_abb_census <- merge(unemployment_abb, census_regions_50, by.x = "Code",by.y = "State")
```

#### Removing the extra state variables - Code and Abbrev
```{r}
unemployment_abb_census <- unemployment_abb_census %>% 
  select(-Code) %>%  select(-Abbrev)
```

#### Making the variable mean_unemployment_rates by subtracting mean unemployment rates in March 2021 by March 2022
```{r}
mean_unemployment_rates <- unemployment_abb_census %>% 
  group_by(Region) %>% 
  summarise(mean= mean(March_21 - March_22))
```

#### Making the variable sd_unemployment_rates by subtracting standard deviation of unemployment rates in March 2022 FROM March 2021
```{r}
sd_unemployment_rates <- unemployment_abb_census %>% 
  group_by(Region) %>% 
  summarise("Standard Deviation "= sd(March_21 - March_22))
```

#### Making the variable cv_unemployment_rates by subtracting coefficient of variation unemployment rates in March 2021 by March 2022
```{r}
cv_unemployment_rates <- unemployment_abb_census %>% 
  group_by(Region) %>% 
  summarise("Coefficient of variation" = sd(March_21 - March_22)/mean(March_21 - March_22))
```

#### Merging the data sets together into "mean_std_cv"
```{r}
mean_std<- merge(mean_unemployment_rates ,sd_unemployment_rates, by = "Region" )

mean_std_cv <- merge(mean_std, cv_unemployment_rates, by = "Region")
```


|           	| Mean 	| Standard Deviation 	| Coefficient of Variation 	|
|-----------	|------	|--------------------	|--------------------------	|
| Midwest   	| 1.34 	| 0.46               	| 0.34                     	|
| Northeast 	| 2.11 	| 0.75               	| 0.35                     	|
| South     	| 1.54 	| 0.49               	| 0.32                     	|
| West      	| 2.15 	| 0.91               	| 0.42                     	|



